### 2017-05-23 doc/categories.tex
More fussing with fig:rsquared and around.

\begin{enumerate}
	\item {\sc LR word}: word type
	\item {\sc LR dep}: word type, dependency label and their interaction 
	\item {\sc LR position}: word type, position (binned as {\sc first, second, third, middle,
	antepenult, penult, last}) and their interaction
	\item {\sc LR full}: word type, dependency label, position, word:dependency interaction, 
	word:position interaction
\end{enumerate}

\noindent We split the whole MSCOCO validation set into two parts, fit the 
models (regularized via L2 penalty) on the first part, and compute the 
proportion of variance explained on the second part. 
For comparison we also train a version of {\sc Imaginet},
where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
over word-embeddings and thus does not have access to sequential cues. We refer to this
model as {\sc Sum}.

Figure~\ref{fig:rsquared} shows the increase in $R^2$  for the models 
per pathway relative to {\sc LR position}. The plot reveals that 
adding additional information on top of lexical features in case of {\sc Sum} 
does increase the explained variance slightly, which is probably due to 
the unseen words in the held out set. More interestingly, the
relative increase in $R^2$ between {\sc LR word} and {\sc LR full} in
case of {\sc Visual} is larger, suggesting that 
the omission scores in case of {\sc Visual} do not only depend on the words themselves,
but their grammatical function and position in sentences.  The raw  $R^2$
scores show that for the {\sc Textual} model the word-type predicts the omission-score
to much smaller degree compared to {\sc Visual} (0.38 against 0.48), and the 
increase in $R^2$ shows that the position of the word adds considerable 
amount of information.

@@

In order to find out some of the specific configurations leading to
increase in $R^2$ in case of {\sc Visual}, we next considered all word types with
occurrence counts of at least 100 and ranked them according to how much
better, on average, {\sc LM deprel} predicted their omission scores 
compared to {\sc LM word}. Figure~\ref{fig:top_words} shows the per-dependency 
omission-score distributions for the five top ranked words plus the word {\it  water}. 
There are clear and large differences in how these words
impact the network's representation depending on what grammatical
function they fulfill. They all have large omission scores when they
occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
grammatical functions typically have a large contribution to the
complete meaning of a sentence.  Conversely, all have small omission 
scores when appearing as {\sc conj} (conjunct): this is probably because in this position
they share their contribution with the first, often more important,
member of the conjunction (e.g. {\it A cow and its baby eating
  grass}).

>>>>

\begin{enumerate}
	\item {\sc LR word}: word type
	\item {\sc LR +dep}: word type, dependency label and their interaction 
	\item {\sc LR +pos}: word type, position (binned as {\sc first, second, third, middle,
	antepenult, penult, last}) and their interaction
	\item {\sc LR full}: word type, dependency label, position, word:dependency interaction, 
	word:position interaction
\end{enumerate}

\noindent We split the 5-thousand-image portion of MSCOCO validation data into two parts, fit the 
models (regularized via L2 penalty) on the first part, and compute the 
proportion of variance explained on the second part. 
For comparison we also train a version of {\sc Imaginet},
where the GRU in the {\sc Visual} pathway is replaced by a simple vector-summation operation
over word-embeddings and thus does not have access to sequential cues. We refer to this
model as {\sc Sum}.

\begin{table}
  \centering
  \caption{Proportion of variance in omission scores explained by
    linear regression.}
    \begin{tabular}{l|rrrr}
               & word   & +pos  & +dep  & full \\\hline
     sum       & 0.654  & 0.661 & 0.670 & 0.670 \\
     LM        & 0.358  & 0.586 & 0.415 & 0.601 \\
     textual   & 0.364  & 0.703 & 0.451 & 0.715 \\
     visual    & 0.490  & 0.506 & 0.515 & 0.523 \\
    \end{tabular}
    \label{tab:lr-r2}
\end{table}

Table~\ref{tab:lr-r2} shows the proportion of variance $R^2$ in omission
scores explained by the linear regression with the different predictors.
Figure~\ref{fig:rsquared} offers a different view of the data, showing
the increase or decrease in $R^2$ for the models per pathway relative to {\sc LR
  +pos}.  The results reveal that adding additional information on
top of lexical features in case of {\sc Sum} does increase the
explained variance slightly, which is probably due to the unseen words
in the held out set. More interestingly, there is a sizeable
increase in $R^2$ between {\sc LR +pos} and {\sc LR full} in
case of {\sc Visual}, suggesting that 
the omission scores in case of {\sc Visual} depend on the words'
grammatical function in sentences, even after controling for word
identity and linear position.

The raw $R^2$ scores show that for the {\sc Textual} and {\sc LM}
models the word-type predicts the omission-score to much smaller
degree compared to {\sc Visual}, and the increase in $R^2$ shows that
the position of the word adds considerable amount of information for
these language models: this is not surprising considering that the omission
scores are measured with respect to the final activation state. 
Less predictably, for the {\sc Visual} model,
dependency labels are more informative than linear position,
hinting at the importance of syntactic structure beyond linear order.

Overall, when regressing on word identities, word position and
dependency labels, the {\sc Visual} model's omission scores are the
hardest to predict of the four models suggesting that the model may be
encoding additional structural features not captured by these
predictors.

@@

In order to find out some of the specific syntactic configurations leading to
increase in $R^2$ between {\sc LR word} and {\sc LR dep} predictors
in case of {\sc Visual}, we next considered all word types with
occurrence counts of at least 100 and ranked them according to how much
better, on average, {\sc LR dep} predicted their omission scores 
compared to {\sc LM +pos}.

Figure~\ref{fig:top_words} shows the per-dependency 
omission-score distributions for the five top ranked words plus the word {\it  water}. 
There are clear and large differences in how these words
impact the network's representation depending on what grammatical
function they fulfill. They all have large omission scores when they
occur as {\sc nsubj} (nominal subject) or {\sc root}, likely due to the fact that these
grammatical functions typically have a large contribution to the
complete meaning of a sentence.  Conversely, all have small omission 
scores when appearing as {\sc conj} (conjunct): this is probably because in this position
they share their contribution with the first, often more important,
member of the conjunction (e.g. {\it A cow and its baby eating
  grass}).
